Calendar-based graphics for visualizing people's daily schedules

### 2017-09-06 main.Rmd
re-working abstract and intro

---
author:
  - name: Earo Wang
    affiliation: Monash University
    address: >
      Department of Econometrics and Business Statistics,
      Monash University, VIC 3800
      Australia
    email: \email{earo.wang@monash.edu}
  - name: Dianne Cook
    affiliation: Monash University
    address: >
      Department of Econometrics and Business Statistics,
      Monash University, VIC 3800
      Australia
    email: \email{dicook@monash.edu}
  - name: Rob J Hyndman
    affiliation: Monash University
    address: >
      Department of Econometrics and Business Statistics,
      Monash University, VIC 3800
      Australia
    email: \email{rob.hyndman@monash.edu}
title:
  formatted: "Calendar-based graphics for visualizing people's daily schedules"
  plain: "Calendar-based graphics for visualizing people's daily schedules"
abstract: >
  This paper describes a \code{frame\_calendar} function that organizes and displays temporal data, collected on sub-daily resolution, into a calendar layout. Calendars are broadly used in society to display temporal information, and events. The \code{frame\_calendar} uses linear algebra on the date variable to create the layout. It utilizes the grammar of graphics to create the plots inside each cell, and thus synchronizes neatly with \pkg{ggplot2} graphics. The motivating application is studying pedestrian behavior in Melbourne, Australia, based on counts which are captured at hourly intervals by sensors scattered around the city. Faceting by the usual features such as day and month, was insufficient to examine the behavior. Making displays on a monthly calendar format helps to understand pedestrian patterns relative to events such as work days, weekends, holidays, and special events. The layout algorithm has several format options and variations. It is implemented in the \proglang{R} package \pkg{sugrrants}.

keywords:
  # at least one keyword must be supplied
  formatted: [data visualization, statistical graphics, time series, grammar of graphics, "\\proglang{R}"]
  plain:     [data visualization, statistical graphics, time series, grammar of graphics, R]
preamble: >
  \usepackage{amsmath}
output: 
  bookdown::pdf_book:
    base_format: rticles::jss_article
  citation_package: biblatex
bibliography: ["references.bib", "packages.bib"]
---

@@

We develop a method for organizing and visualizing temporal data, collected at sub-daily intervals, into a calendar layout. The calendar format is created using linear algebra, and fits into the \pkg{tidyverse} [@R-tidyverse] and grammar of graphics [@wickham2009ggplot2] framework for exploring data. The purpose of the calendar-based visualization is to provide insights into people's daily schedules relative to events such as work days, weekends, holidays, and special events. This work was originally motivated by studying foot traffic in the city of Melbourne, Australia [@ped]. There have been 43 sensors installed that count pedestrians every hour across the downtown area (Figure \ref{fig:ped-map}). The dataset can shed light on people's daily rhythms, and assist the city administration and local businesses with event planning and operational management. A routine examination of the data would involve constructing conventional time series plots to catch a glimpse of temporal patterns. The faceted plots in Figure \ref{fig:time-series-plot}, give an overall picture of the foot traffic at 3 different sensors over 2016. Further faceting by day of the week (Figure \ref{fig:facet-time}) provides a better glimpse of the daily and sub-daily pedestrian patterns. The sensor data, like many temporal datasets on human behaviors, lends itself to a number of exploratory data visualization challenges:

>>>>

---
author:
  - name: Earo Wang
    affiliation: Monash University
    address: >
      Department of Econometrics and Business Statistics,
      Monash University, VIC 3800
      Australia
    email: \email{earo.wang@monash.edu}
  - name: Dianne Cook
    affiliation: Monash University
    address: >
      Department of Econometrics and Business Statistics,
      Monash University, VIC 3800
      Australia
    email: \email{dicook@monash.edu}
  - name: Rob J Hyndman
    affiliation: Monash University
    address: >
      Department of Econometrics and Business Statistics,
      Monash University, VIC 3800
      Australia
    email: \email{rob.hyndman@monash.edu}
title:
  formatted: "Calendar-based graphics for visualizing people's daily schedules"
  plain: "Calendar-based graphics for visualizing people's daily schedules"
abstract: >
  Calendars are broadly used in society to display temporal information, and events. This paper describes a new \proglang{R} package with functionality to organize and display temporal data, collected on sub-daily resolution, into a calendar layout. The function \code{frame\_calendar} uses linear algebra on the date variable to re-structure data into a format lending itself to calendar layouts. The user can apply the grammar of graphics to create plots inside each calendar cell, and thus the displays synchronize neatly with \pkg{ggplot2} graphics. The motivating application is studying pedestrian behavior in Melbourne, Australia, based on counts which are captured at hourly intervals by sensors scattered around the city. Faceting by the usual features such as day and month, was insufficient to examine the behavior. Making displays on a monthly calendar format helps to understand pedestrian patterns relative to events such as work days, weekends, holidays, and special events. The layout algorithm has several format options and variations. It is implemented in the \proglang{R} package \pkg{sugrrants}.
keywords:
  # at least one keyword must be supplied
  formatted: [data visualization, statistical graphics, time series, grammar of graphics, "\\proglang{R}"]
  plain:     [data visualization, statistical graphics, time series, grammar of graphics, R]
preamble: >
  \usepackage{amsmath}
output: 
  bookdown::pdf_book:
    base_format: rticles::jss_article
  citation_package: biblatex
bibliography: ["references.bib", "packages.bib"]
---

@@

We develop a method for organizing and visualizing temporal data, collected at sub-daily intervals, into a calendar layout. The calendar format is created using linear algebra, giving a restructuring of the data, that can then be piped into grammar of graphics definitions of plots, as used in \pkg{ggplot2} [@R-ggplot2]. The data restructuring approach is consistent with the tidy data principles available in the \pkg{tidyverse} [@R-tidyverse] suite. The methods are implemented in a new package called \pkg{sugrrants}.

The purpose of the calendar-based visualization is to provide insights into people's daily schedules relative to events such as work days, weekends, holidays, and special events. This work was originally motivated by studying foot traffic in the city of Melbourne, Australia [@ped]. There have been 43 sensors installed that count pedestrians every hour across the downtown area (Figure \ref{fig:ped-map}). The dataset can shed light on people's daily rhythms, and assist the city administration and local businesses with event planning and operational management. A routine examination of the data would involve constructing conventional time series plots to catch a glimpse of temporal patterns. The faceted plots in Figure \ref{fig:time-series-plot}, give an overall picture of the foot traffic at 3 different sensors over 2016. Further faceting by day of the week (Figure \ref{fig:facet-time}) provides a better glimpse of the daily and sub-daily pedestrian patterns. The sensor data, like many temporal datasets on human behaviors, lends itself to a number of exploratory data visualization challenges:

### 2017-09-06 main.tex
re-working abstract and intro

\Abstract{
This paper describes a \code{frame\_calendar} function that organizes
and displays temporal data, collected on sub-daily resolution, into a
calendar layout. Calendars are broadly used in society to display
temporal information, and events. The \code{frame\_calendar} uses linear
algebra on the date variable to create the layout. It utilizes the
grammar of graphics to create the plots inside each cell, and thus
synchronizes neatly with \pkg{ggplot2} graphics. The motivating
application is studying pedestrian behavior in Melbourne, Australia,
based on counts which are captured at hourly intervals by sensors
scattered around the city. Faceting by the usual features such as day
and month, was insufficient to examine the behavior. Making displays on
a monthly calendar format helps to understand pedestrian patterns
relative to events such as work days, weekends, holidays, and special
events. The layout algorithm has several format options and variations.
It is implemented in the \proglang{R} package \pkg{sugrrants}.
}

@@

\usepackage{amsthm}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}{Lemma}[section]
\theoremstyle{definition}
\newtheorem{definition}{Definition}[section]
\newtheorem{corollary}{Corollary}[section]
\newtheorem{proposition}{Proposition}[section]
\theoremstyle{definition}
\newtheorem{example}{Example}[section]
\theoremstyle{definition}
\newtheorem{exercise}{Exercise}[section]
\theoremstyle{remark}
\newtheorem*{remark}{Remark}
\newtheorem*{solution}{Solution}
\begin{document}

@@

We develop a method for organizing and visualizing temporal data,
collected at sub-daily intervals, into a calendar layout. The calendar
format is created using linear algebra, and fits into the
\pkg{tidyverse} \citep{R-tidyverse} and grammar of graphics
\citep{wickham2009ggplot2} framework for exploring data. The purpose of
the calendar-based visualization is to provide insights into people's
daily schedules relative to events such as work days, weekends,
holidays, and special events. This work was originally motivated by
studying foot traffic in the city of Melbourne, Australia \citep{ped}.
There have been 43 sensors installed that count pedestrians every hour
across the downtown area (Figure \ref{fig:ped-map}). The dataset can
shed light on people's daily rhythms, and assist the city administration
and local businesses with event planning and operational management. A
routine examination of the data would involve constructing conventional
time series plots to catch a glimpse of temporal patterns. The faceted
plots in Figure \ref{fig:time-series-plot}, give an overall picture of
the foot traffic at 3 different sensors over 2016. Further faceting by
day of the week (Figure \ref{fig:facet-time}) provides a better glimpse
of the daily and sub-daily pedestrian patterns. The sensor data, like
many temporal datasets on human behaviors, lends itself to a number of
exploratory data visualization challenges:

>>>>

\Abstract{
Calendars are broadly used in society to display temporal information,
and events. This paper describes a new \proglang{R} package with
functionality to organize and display temporal data, collected on
sub-daily resolution, into a calendar layout. The function
\code{frame\_calendar} uses linear algebra on the date variable to
re-structure data into a format lending itself to calendar layouts. The
user can apply the grammar of graphics to create plots inside each
calendar cell, and thus the displays synchronize neatly with
\pkg{ggplot2} graphics. The motivating application is studying
pedestrian behavior in Melbourne, Australia, based on counts which are
captured at hourly intervals by sensors scattered around the city.
Faceting by the usual features such as day and month, was insufficient
to examine the behavior. Making displays on a monthly calendar format
helps to understand pedestrian patterns relative to events such as work
days, weekends, holidays, and special events. The layout algorithm has
several format options and variations. It is implemented in the
\proglang{R} package \pkg{sugrrants}.
}

@@

\usepackage{amsthm}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}{Lemma}[section]
\theoremstyle{definition}
\newtheorem{definition}{Definition}[section]
\newtheorem{corollary}{Corollary}[section]
\newtheorem{proposition}{Proposition}[section]
\theoremstyle{definition}
\newtheorem{example}{Example}[section]
\theoremstyle{remark}
\newtheorem*{remark}{Remark}
\begin{document}

@@

We develop a method for organizing and visualizing temporal data,
collected at sub-daily intervals, into a calendar layout. The calendar
format is created using linear algebra, giving a restructuring of the
data, that can then be piped into grammar of graphics definitions of
plots, as used in \pkg{ggplot2} \citep{R-ggplot2}. The data
restructuring approach is consistent with the tidy data principles
available in the \pkg{tidyverse} \citep{R-tidyverse} suite. The methods
are implemented in a new package called \pkg{sugrrants}.

The purpose of the calendar-based visualization is to provide insights
into people's daily schedules relative to events such as work days,
weekends, holidays, and special events. This work was originally
motivated by studying foot traffic in the city of Melbourne, Australia
\citep{ped}. There have been 43 sensors installed that count pedestrians
every hour across the downtown area (Figure \ref{fig:ped-map}). The
dataset can shed light on people's daily rhythms, and assist the city
administration and local businesses with event planning and operational
management. A routine examination of the data would involve constructing
conventional time series plots to catch a glimpse of temporal patterns.
The faceted plots in Figure \ref{fig:time-series-plot}, give an overall
picture of the foot traffic at 3 different sensors over 2016. Further
faceting by day of the week (Figure \ref{fig:facet-time}) provides a
better glimpse of the daily and sub-daily pedestrian patterns. The
sensor data, like many temporal datasets on human behaviors, lends
itself to a number of exploratory data visualization challenges:
